---
title: "RiparianBuffers"
author: "Shirin Varkouhi"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE, message=FALSE}
library(SpaDES.core)
library(SpaDES.project)
library(terra)
library(sf)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Overview

**RiparianBuffers** computes a spatially explicit fractional riparian influence raster (0–1)  
based on hydrological features and jurisdiction-specific buffer policies.

The module is:

- Policy-aware  
- Decision-neutral  
- Spatial-signal generating only  

It does **not** apply regulatory exclusions or define landbase decisions.

---

# Conceptual Role in the Model

This module translates:

Hydrology × Jurisdiction × Buffer Policy  

into  

A continuous riparian influence surface.

The output is intended for use by downstream modules  
(e.g., landbase, AAC, biodiversity, habitat models).

---

# What This Module Does

- Computes fractional riparian influence (0–1)
- Uses streams and lakes as hydrological drivers
- Applies jurisdiction-specific buffer widths
- Produces a reusable spatial signal
- Supports standalone execution (downloads HydroRIVERS/HydroLAKES if needed)

---

# What This Module Does Not Do

- Does not define harvestable landbase
- Does not apply regulatory exclusions
- Does not interpret legal compliance
- Does not classify forest/non-forest
- Does not make management decisions

---

# Required Inputs

The module expects the following objects in `sim`:

## 1. PlanningGrid_250m (`SpatRaster`)
Coarse-resolution planning grid.

## 2. studyArea (`SpatVector` or `sf`)
Polygon used for cropping hydrological inputs.

## 3. riparianBufferPolicy (`data.frame`)
Table defining buffer widths by jurisdiction and hydrology class.

## 4. Hydrology_streams (`SpatVector`) (optional)
Stream network (e.g., HydroRIVERS).

## 5. Hydrology_lakes (`SpatVector`) (optional)
Lake polygons (e.g., HydroLAKES).

If hydrology inputs are not supplied upstream,  
the module automatically downloads and prepares them.

---

# Jurisdiction Layer

If a jurisdiction layer is not provided,  
the module builds one internally using `rnaturalearth`.

Provinces included:

- Ontario (ON)
- Quebec (QC)
- New Brunswick (NB)
- Nova Scotia (NS)
- Prince Edward Island (PE)
- Newfoundland and Labrador (NL)

The layer is:

- Cropped to `studyArea`
- Projected to match the planning grid
- Stored as `sim$jurisdiction`

---

# Buffer Policy Structure

Buffer widths must be provided in the following format:

```{r policy_example, eval=FALSE}
data.frame(
  jurisdiction = c("default", "ON", "QC"),
  small_stream = c(30, 30, 20),
  large_stream = c(50, 50, 40),
  small_lake   = c(30, 30, 20),
  large_lake   = c(50, 50, 40)
)
```

Rules:

- Distances are interpreted geometrically only.
- No regulatory interpretation occurs here.
- If a jurisdiction-specific rule is missing, `"default"` is used.

---

# Computational Resolution

Riparian influence is computed at a higher resolution defined by:

`hydroRaster_m`

Default: 30 m

This allows narrow hydrological features to be captured  
independently from the 250 m planning grid.

The fractional influence is then aggregated back to `PlanningGrid_250m`.

---

# Output Object

The module creates:

## sim$Riparian (list)

Contains:

- `riparianFraction` – SpatRaster (0–1)
- `raster_m` – computation resolution
- `policy` – buffer policy table used

---

# Interpretation of Output

| Value | Meaning |
|-------|---------|
| 0     | No riparian influence |
| 1     | Fully riparian |
| 0–1   | Partial influence |

Interpretation and decision rules are handled downstream.

---

# Example Usage

```{r module_usage, eval=FALSE}
times <- list(start = 0, end = 1)

modules <- list("RiparianBuffers")

parameters <- list(
  RiparianBuffers = list(
    hydroRaster_m = 30
  )
)

objects <- list(
  PlanningGrid_250m    = PlanningGrid_250m,
  studyArea            = studyArea,
  Hydrology_streams    = Hydrology_streams,
  Hydrology_lakes      = Hydrology_lakes,
  riparianBufferPolicy = riparianBufferPolicy
)

sim <- simInit(
  times   = times,
  modules = modules,
  params  = parameters,
  objects = objects
)

sim <- spades(sim)
```

---

# Anticipated Linkages

RiparianBuffers is designed to precede:

- Landbase estimation modules
- Harvest/AAC modules
- Habitat and biodiversity modules
- Spatial weighting or masking frameworks

---

# Plotting

None.

---

# Saving

None.

All outputs remain in-memory within the simulation object.