---
title: "RiparianBuffers"
author: "Shirin Varkouhi"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
---
```{r setup, include=TRUE, echo=FALSE, message=FALSE}
library(SpaDES.core)
library(SpaDES.project)
library(terra)
library(sf)
knitr::opts_chunk$set(echo = TRUE)
```
## Overview

This module computes a fractional riparian influence raster (values between 0 and 1)
from hydrological features and jurisdiction-specific buffer rules.

The module is policy-aware but decision-free:
it produces a spatial signal only and leaves all regulatory or management interpretation
to downstream modules.

## Plotting

Nothing.

## Saving

Nothing.

All outputs are stored in memory as part of the simulation object.

## What this module does

- Computes fractional riparian influence (0–1)
- Uses streams and lakes as hydrological features
- Supports jurisdiction-specific buffer widths
- Produces outputs reusable by downstream modules

## What this module does not do

- Does not define harvestable landbase
- Does not apply regulatory exclusions
- Does not interpret policy compliance
- Does not classify forest vs non-forest

## Input data

This module expects the following input objects:

- **PlanningRaster** (`SpatRaster`)  
  Coarse-resolution planning grid

- **Hydrology_streams** (`SpatVector`)  
  Stream and river network (e.g. HydroRIVERS)

- **Hydrology_lakes** (`SpatVector`)  
  Lakes and large water bodies (e.g. HydroLAKES)

- **Provinces** (`SpatVector`)  
  Jurisdiction boundaries with a `jurisdiction` field

- **riparianBufferPolicy** (`data.frame`)  
  Table defining buffer widths by jurisdiction and hydrology class
  
## Study area

The study area is supplied upstream as spatial input objects
and is not defined internally in this module.

All hydrological and jurisdictional inputs are expected
to be pre-cropped and projected consistently
(e.g. by EasternCanadaDataPrep).

## Buffer policy

Riparian buffer widths are supplied as a data frame.

Example:

```{r policy_example, eval=FALSE}
data.frame(
  jurisdiction = c("default", "ON", "QC"),
  small_stream = c(30, 30, 20),
  large_stream = c(50, 50, 40),
  small_lake   = c(30, 30, 20),
  large_lake   = c(50, 50, 40)
)
```
Buffer distances are applied geometrically only.
No regulatory interpretation is performed in this module.

---

## Example usage

```{r module_usage, message=FALSE, eval=FALSE}
times <- list(start = 0, end = 1)

modules <- list("RiparianBuffers")

parameters <- list(
  RiparianBuffers = list(
    hydroRaster_m = 30
  )
)

objects <- list(
  PlanningRaster    = PlanningRaster,
  Hydrology_streams = Hydrology_streams,
  Hydrology_lakes   = Hydrology_lakes,
  Provinces         = Provinces
)

sim <- simInit(
  times   = times,
  modules = modules,
  params  = parameters,
  objects = objects
)

sim <- spades(sim)
```


---

## Output data

The module creates a single output object:

- **sim$Riparian**
  - `riparianFraction` – SpatRaster (0–1)
  - `raster_m` – computation resolution (m)
  - `policy` – buffer policy table used

## Interpretation

- 0 indicates no riparian influence
- 1 indicates full riparian coverage
- Intermediate values indicate partial influence

Downstream modules determine how this signal is used
(e.g. masking, weighting, reporting).
## Anticipated linkages to other modules

RiparianBuffers is designed to be used upstream of:

- landbase estimation modules
- harvesting and AAC modules
- habitat and biodiversity analyses
